---
title: "Eukaryotic SP analysis by taxonomy"
author: "Katarzyna Sidorczuk"
date: "18/08/2022"
output: 
  html_document:
  toc: true
params:
  encoding: NULL
  k: 2
  gaps: 0
  fcbf_amp: FALSE
  quipt_amp: TRUE
  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.width = 14, fig.height = 10, warning = FALSE, message = FALSE, cache = TRUE, cache.lazy = FALSE)

library(targets)
library(dplyr)
library(ggplot2)
library(seqR)
library(tidyr)
library(pbapply)
library(stringr)
library(ggnewscale)
library(biogram)
library(FCBF)
library(DT)
library(ggVennDiagram)
library(ggdendro)
library(grid)
library(gridExtra)

source("./functions/extract_features.R")
source("./functions/motif_functions.R")
tar_load(c(datasets_list, taxonomic_representation))

my_DT <- function(df) {
  datatable(df, extensions = 'FixedColumns', filter = "top", style = "bootstrap",
            options = list(dom = 't', scrollX = TRUE, scrollCollapse = TRUE,
                           scrollY = "400px", paging = FALSE), rownames = FALSE) %>% 
    formatRound(2:ncol(df), 4) 
}

colors <- c("cTP experimentally verified presequence" = "#9de444", "#e4d345", "#528313", "#837713",
            "mTP experimentally verified presequence" = "#e49144", "#834813", "SP experimentally verified presequence" = "#45e495", 
            "#13834b", "#44aee4", "#4451e4", "#e44444", "#831313", "#e44496", "#bb45e4", "#8d44e4", "#83134c", "#661383", "#461383")
colors2 <- c("#c4e7d4", "#c4dacf", "#bfcdd5", "#b9c0da", "#b1b4cc", "#a9a7bd", "#a19aaf", "#9d94a8", "#998da0", "#63585e",
             "#270722", "#402030", "#59393d", "#72524b", "#8a6b58", "#ecce8e", "#dbcf96", "#cfcb9f", "#c2c6a7", "#d6d8c3",
             "#333333", "#4d4f5d", "#666a86", "#727e99", "#7e91ac", "#8aa5bf", "#95b8d1", "#bfcbc3", "#d4d4bc", "#e8ddb5",
             "#efd0ca", "#c1bcac", "#979b8d", "#7a8872", "#5c7457")

filtered_datasets_list <- datasets_list[c("cTP experimentally verified presequence", 
                                          "mTP experimentally verified presequence", 
                                          "SP experimentally verified presequence")]

filtered_datasets <- unlist(sapply(names(filtered_datasets_list), function(i) rep(i, length(filtered_datasets_list[[i]])), USE.NAMES = FALSE))

```


# SP
```{r}
ds <- filtered_datasets_list[["SP experimentally verified presequence"]]
df <- filter(taxonomic_representation, Entry %in% names(ds))
df[is.na(df)] <- "NA"
```

## By kingdom (Viridiplantae, Metazoa, Fungi)
```{r}
SP_kingdom <- lapply(c("Viridiplantae", "Metazoa", "Fungi"), function(i) {
  ds[which(names(ds) %in% filter(df, kingdom == i)[["Entry"]])]
}) %>% setNames(c("Viridiplantae", "Metazoa", "Fungi"))

lengths(SP_kingdom) %>% 
  data.frame() %>% 
  setNames("Number of sequences") %>% 
  knitr::kable()

trigram_freqs <- calculate_ngram_freq(SP_kingdom, params[["k"]], params[["gaps"]], params[["encoding"]])
trigram_plot_dat_presequences <- get_motif_plot_dat(trigram_freqs)
n_d <- 3
```

### Motifs occuring in all kingdoms
```{r fig.height = 20}
motif_counts_datasets2 <- trigram_plot_dat_presequences %>% 
  pivot_wider(motif, names_from = dataset, values_from = frequency) %>% 
  mutate(n_datasets = apply(select(., -motif), 1, function(r) sum(r!=0)))

if(nrow(filter(motif_counts_datasets2, n_datasets == n_d)) > 0) {
  filter(motif_counts_datasets2, n_datasets == n_d) %>% 
    select(-n_datasets) %>% 
    pivot_longer(2:ncol(.), names_to = "Kingdom", values_to = "frequency") %>% 
    ggplot(aes(x = Kingdom, y = frequency, fill = Kingdom)) +
    geom_col() +
    facet_wrap(~motif, ncol = 20) +
    theme_bw() +
    theme(legend.position = "bottom",
          axis.text.x = element_blank(),
          axis.ticks.x = element_blank()) 
  
  filter(motif_counts_datasets2, n_datasets == n_d) %>% 
    select(-n_datasets) %>% 
    my_DT()
}
```

### Motifs unique for specific kingdom

```{r}
x <- filter(motif_counts_datasets2, n_datasets == 1) %>% 
  select(-n_datasets) 

if(length(x) > 0) {
  pivoted_x <- pivot_longer(x, 2:ncol(x), names_to = "Kingdom", values_to = "frequency") %>% 
    filter(frequency > 0) 
  
  group_by(pivoted_x, Kingdom) %>% 
    summarise(`Number of unique motifs` = n()) %>% 
    my_DT()
}

```


### The most frequent motifs for kingdoms
```{r results = 'asis'}
for(i in names(SP_kingdom)) {
  most_frequent <- get_dataset_most_frequent_motifs(trigram_plot_dat_presequences,
                                                    i,
                                                    n_motifs = 100)
  
  plot(plot_most_frequent_motifs(trigram_plot_dat_presequences,
                                 most_frequent,
                                 unname(colors),
                                 paste0("100 most frequent motifs of ", i)))
  
  print(filter(trigram_plot_dat_presequences, motif %in% most_frequent) %>% 
          pivot_wider(names_from = "dataset", values_from = "frequency") %>% 
          my_DT() %>% 
          htmltools::tagList())
}
```

### Position density of the 50 most frequent motifs for each kingdom
```{r, results = 'asis'}
for(i in names(SP_kingdom)) {
  most_frequent <- get_dataset_most_frequent_motifs(trigram_plot_dat_presequences,
                                                    i,
                                                    n_motifs = 50)
  plot(plot_motif_position_density(SP_kingdom, i,
                                   locate_motifs(SP_kingdom[i], most_frequent, params[["encoding"]]),
                                   density_col = colors[["SP experimentally verified presequence"]]))
}
```

### Motif positions within scaled sequences
```{r}
for(i in names(SP_kingdom)) {
  most_frequent <- get_dataset_most_frequent_motifs(trigram_plot_dat_presequences,
                                                    i,
                                                    n_motifs = 50)
  plot(plot_motif_positions_scaled(SP_kingdom, i,
                                   locate_motifs(SP_kingdom[i], most_frequent, params[["encoding"]]),
                                   type = "violin", density_col = colors[["SP experimentally verified presequence"]]))
}
```


## Metazoa by phylum (Arthropoda, Chordata)
```{r}
SP_phylum <- lapply(c("Arthropoda", "Chordata"), function(i) {
  ds[which(names(ds) %in% filter(df, phylum == i)[["Entry"]])]
}) %>% setNames(c("Arthropoda", "Chordata"))

lengths(SP_phylum) %>%
  data.frame() %>%
  setNames("Number of sequences") %>%
  knitr::kable()

trigram_freqs <- calculate_ngram_freq(SP_phylum, params[["k"]], params[["gaps"]], params[["encoding"]])
trigram_plot_dat_presequences <- get_motif_plot_dat(trigram_freqs)
n_d <- length(names(SP_phylum))
```

### Motifs occuring in all phyla
```{r}
motif_counts_datasets2 <- trigram_plot_dat_presequences %>%
  pivot_wider(motif, names_from = dataset, values_from = frequency) %>%
  mutate(n_datasets = apply(select(., -motif), 1, function(r) sum(r!=0)))

if(nrow(filter(motif_counts_datasets2, n_datasets == n_d)) > 0) {
  filter(motif_counts_datasets2, n_datasets == n_d) %>%
    select(-n_datasets) %>%
    pivot_longer(2:ncol(.), names_to = "Phylum", values_to = "frequency") %>%
    ggplot(aes(x = Phylum, y = frequency, fill = Phylum)) +
    geom_col() +
    facet_wrap(~motif, ncol = 20) +
    theme_bw() +
    theme(legend.position = "bottom",
          axis.text.x = element_blank(),
          axis.ticks.x = element_blank())
  
  filter(motif_counts_datasets2, n_datasets == n_d) %>%
    select(-n_datasets) %>%
    my_DT() 
}
```

### Motifs unique for specific phyla
```{r}
x <- filter(motif_counts_datasets2, n_datasets == 1) %>%
  select(-n_datasets)

if(length(x) > 0) {
  pivoted_x <- pivot_longer(x, 2:ncol(x), names_to = "Phylum", values_to = "frequency") %>%
    filter(frequency > 0)
  
  group_by(pivoted_x, Phylum) %>%
    summarise(`Number of unique motifs` = n()) %>%
    my_DT() 
}
```


### The most frequent motifs for phyla
```{r results = 'asis'} 
for(i in names(SP_phylum)) {
  most_frequent <- get_dataset_most_frequent_motifs(trigram_plot_dat_presequences,
                                                    i,
                                                    n_motifs = 100)
  plot(plot_most_frequent_motifs(trigram_plot_dat_presequences,
                                 most_frequent,
                                 unname(colors),
                                 paste0("100 most frequent motifs of ", i)))
  
  print(filter(trigram_plot_dat_presequences, motif %in% most_frequent) %>%
          pivot_wider(names_from = "dataset", values_from = "frequency") %>%
          my_DT() %>% 
          htmltools::tagList())
}

```



## Metazoa by class (Arachnida, Insecta, Lepidosauria, Mammalia)
```{r}
SP_class <- lapply(c("Arachnida", "Insecta", "Lepidosauria", "Mammalia"), function(i) {
  ds[which(names(ds) %in% filter(df, class == i)[["Entry"]])]
}) %>% setNames(c("Arachnida", "Insecta", "Lepidosauria", "Mammalia"))

lengths(SP_class) %>%
  data.frame() %>%
  setNames("Number of sequences") %>%
  knitr::kable()

trigram_freqs <- calculate_ngram_freq(SP_class, params[["k"]], params[["gaps"]], params[["encoding"]])
trigram_plot_dat_presequences <- get_motif_plot_dat(trigram_freqs)
n_d <- length(names(SP_class))
```

### Motifs occuring in all classes
```{r}
motif_counts_datasets2 <- trigram_plot_dat_presequences %>%
  pivot_wider(motif, names_from = dataset, values_from = frequency) %>%
  mutate(n_datasets = apply(select(., -motif), 1, function(r) sum(r!=0)))

if(nrow(filter(motif_counts_datasets2, n_datasets == n_d)) > 0) {
  filter(motif_counts_datasets2, n_datasets == n_d) %>%
    select(-n_datasets) %>%
    pivot_longer(2:ncol(.), names_to = "Class", values_to = "frequency") %>%
    ggplot(aes(x = Class, y = frequency, fill = Class)) +
    geom_col() +
    facet_wrap(~motif, ncol = 20) +
    theme_bw() +
    theme(legend.position = "bottom",
          axis.text.x = element_blank(),
          axis.ticks.x = element_blank())
  
  filter(motif_counts_datasets2, n_datasets == n_d) %>%
    select(-n_datasets) %>%
    my_DT() 
}
```

### Motifs unique for specific classes
```{r}
x <- filter(motif_counts_datasets2, n_datasets == 1) %>%
  select(-n_datasets)

if(length(x) > 0) {
  pivoted_x <- pivot_longer(x, 2:ncol(x), names_to = "Class", values_to = "frequency") %>%
    filter(frequency > 0)
  
  group_by(pivoted_x, Class) %>%
    summarise(`Number of unique motifs` = n()) %>%
    my_DT() 
}
```


### The most frequent motifs for classes
```{r results = 'asis'} 
for(i in names(SP_class)) {
  most_frequent <- get_dataset_most_frequent_motifs(trigram_plot_dat_presequences,
                                                    i,
                                                    n_motifs = 100)
  plot(plot_most_frequent_motifs(trigram_plot_dat_presequences,
                                 most_frequent,
                                 colors2,
                                 paste0("100 most frequent motifs of ", i)))
  
  print(filter(trigram_plot_dat_presequences, motif %in% most_frequent) %>%
          pivot_wider(names_from = "dataset", values_from = "frequency") %>%
          my_DT() %>% 
          htmltools::tagList())
}
```

### Position density of the 50 most frequent motifs for each class
```{r results = 'asis'}
for(i in names(SP_class)) {
  most_frequent <- get_dataset_most_frequent_motifs(trigram_plot_dat_presequences,
                                                    i,
                                                    n_motifs = 50)
  plot(plot_motif_position_density(SP_class, i,
                                   locate_motifs(SP_class[i], most_frequent, params[["encoding"]]),
                                   density_col = colors[["SP experimentally verified presequence"]]))
}

```

### Motif positions within scaled sequences
```{r results = 'asis'}
for(i in names(SP_class)) {
  most_frequent <- get_dataset_most_frequent_motifs(trigram_plot_dat_presequences,
                                                    i,
                                                    n_motifs = 50)
  plot(plot_motif_positions_scaled(SP_class, i,
                                   locate_motifs(SP_class[i], most_frequent, params[["encoding"]]),
                                   type = "violin", density_col = colors[["SP experimentally verified presequence"]]))
}
```
