---
title: "Presequences: ngram analysis with encoding"
author: "Katarzyna Sidorczuk"
date: "02/06/2022"
output: 
  html_document:
  toc: true
params:
  encoding: NULL
  k: 2
  gaps: 0
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.width = 14, fig.height = 10, warning = FALSE, message = FALSE, cache = TRUE, cache.lazy = FALSE)

library(targets)
library(dplyr)
library(ggplot2)
library(seqR)
library(tidyr)
library(pbapply)
library(stringr)
library(ggnewscale)
library(biogram)
library(FCBF)
library(DT)
library(ggVennDiagram)
library(ggdendro)
library(grid)
library(gridExtra)

source("./functions/extract_features.R")
source("./functions/motif_functions.R")
withr::with_dir("./", 
                tar_load(datasets_list))

my_DT <- function(df) {
  datatable(df, extensions = 'FixedColumns', filter = "top", style = "bootstrap",
            options = list(dom = 't', scrollX = TRUE, scrollCollapse = TRUE,
                           scrollY = "700px", paging = FALSE), rownames = FALSE)
}

colors <- c("AmyPro regions" = "#e44496", "CPAD peptides" = "#8d44e4", 
            "cTP-mTP experimentally verified presequence" ="#e4d345", 
            "cTP experimentally verified presequence" = "#9de444", "DBAASP AMP max 100 aa" = "#e44444", 
            "mTP experimentally verified presequence" = "#e49144", "SP experimentally verified presequence" = "#45e495", 
            "TM regions experimentally verified - alpha" = "#44aee4", "TM regions experimentally verified - beta" = "#4451e4")

colors2 <- c("cTP-mTP experimentally verified presequence" ="#e4d345", 
             "cTP experimentally verified presequence" = "#9de444", 
             "mTP experimentally verified presequence" = "#e49144", 
             "SP experimentally verified presequence" = "#45e495")

colors3 <- c("cTP-mTP experimentally verified presequence" ="#e4d345", 
             "cTP experimentally verified presequence" = "#9de444", 
             "mTP experimentally verified presequence" = "#e49144", 
             "SP experimentally verified presequence" = "#45e495", 
             "DBAASP AMP max 100 aa" = "#e44444")


filtered_datasets_list <- datasets_list[c("AmyPro regions", "CPAD peptides", "cTP-mTP experimentally verified presequence", 
                                          "cTP experimentally verified presequence", "DBAASP AMP max 100 aa", 
                                          "mTP experimentally verified presequence", "SP experimentally verified presequence", 
                                          "TM regions experimentally verified - alpha", "TM regions experimentally verified - beta")]

filtered_datasets <- unlist(sapply(names(filtered_datasets_list), function(i) rep(i, length(filtered_datasets_list[[i]])), USE.NAMES = FALSE))

trigram_freqs1 <- calculate_ngram_freq(filtered_datasets_list, params[["k"]], params[["gaps"]], params[["encoding"]])
trigram_plot_dat <- get_motif_plot_dat(trigram_freqs1)

```

# Used encoding
```{r}
knitr::kable(data.frame("group" = names(params[["encoding"]]), "aa" = sapply(params[["encoding"]], function(i) paste0(i, collapse = ", "))), row.names = FALSE)
```

# Overall analysis of all data sets

## Motif frequency

### Numbers of distinct motifs
```{r fig.width = 8, fig.height = 3}
trigram_plot_dat %>% 
  filter(frequency > 0) %>% 
  group_by(dataset) %>% 
  summarise(n_motifs = n()) %>% 
  ggplot(aes(x = dataset, y = n_motifs, fill = dataset)) +
  geom_col(color = "black", alpha = 0.5) +
  coord_flip() + 
  theme_bw() +
  scale_fill_manual("Data set", values = colors) +
  theme(legend.position = "none")
```

### Motifs occuring in all datasets
```{r fig.height = 14}
motif_counts_datasets <- trigram_plot_dat %>% 
  pivot_wider(motif, names_from = dataset, values_from = frequency) %>% 
  mutate(n_datasets = apply(select(., -motif), 1, function(r) sum(r!=0)))

print(paste0("Number of motifs occuring in all datasets: ", nrow(filter(motif_counts_datasets, n_datasets == 9))))
if(nrow(filter(motif_counts_datasets, n_datasets == 9)) > 0) {
  filter(motif_counts_datasets, n_datasets == 9) %>% 
    select(-n_datasets) %>% 
    pivot_longer(2:10, names_to = "dataset", values_to = "frequency") %>% 
    ggplot(aes(x = dataset, y = frequency, fill = dataset)) +
    geom_col() +
    facet_wrap(~motif) +
    theme_bw() +
    theme(legend.position = "bottom",
          axis.text.x = element_blank(),
          axis.ticks.x = element_blank()) +
    scale_fill_manual("Data set", values = colors)
}
```


# Detailed analysis of presequences

```{r}
trigram_plot_dat_presequences <- trigram_plot_dat %>% 
  filter(dataset %in% c("cTP-mTP experimentally verified presequence", "cTP experimentally verified presequence", 
                        "mTP experimentally verified presequence", "SP experimentally verified presequence"))

plot_motif_venn_diagram(trigram_plot_dat_presequences, colors2)

```

### Motifs occuring in all presequence datasets
```{r fig.height = 20}
motif_counts_datasets2 <- trigram_plot_dat_presequences %>% 
  pivot_wider(motif, names_from = dataset, values_from = frequency) %>% 
  mutate(n_datasets = apply(select(., -motif), 1, function(r) sum(r!=0)))

if(nrow(filter(motif_counts_datasets2, n_datasets == 4)) > 0) {
  filter(motif_counts_datasets2, n_datasets == 4) %>% 
    select(-n_datasets) %>% 
    pivot_longer(2:5, names_to = "dataset", values_to = "frequency") %>% 
    ggplot(aes(x = dataset, y = frequency, fill = dataset)) +
    geom_col() +
    facet_wrap(~motif, ncol = 20) +
    theme_bw() +
    theme(legend.position = "bottom",
          axis.text.x = element_blank(),
          axis.ticks.x = element_blank()) +
    scale_fill_manual("Data set", values = colors2)
  
  filter(motif_counts_datasets2, n_datasets == 4) %>% 
    select(-n_datasets) %>% 
    my_DT()
}
```

### Motifs unique for specific classes

```{r}
x <- filter(motif_counts_datasets2, n_datasets == 1) %>% 
  select(-n_datasets) 

if(length(x) > 0) {
  pivoted_x <- pivot_longer(x, 2:5, names_to = "dataset", values_to = "frequency") %>% 
    filter(frequency > 0) 
  
  group_by(pivoted_x, dataset) %>% 
    summarise(`Number of unique motifs` = n()) %>% 
    my_DT()
}

```

### Motifs specific for two classes
```{r}
x <- filter(motif_counts_datasets2, n_datasets == 2) %>% 
  select(-n_datasets) 

if(nrow(x) > 0) {
  my_DT(x)
}
```

### The most frequent motifs for specific classes
```{r}
most_frequent_cTP <- get_dataset_most_frequent_motifs(trigram_plot_dat_presequences,
                                                      "cTP experimentally verified presequence",
                                                      n_motifs = 100)

most_frequent_mTP <- get_dataset_most_frequent_motifs(trigram_plot_dat_presequences,
                                                      "mTP experimentally verified presequence",
                                                      n_motifs = 100)

most_frequent_cTPmTP <- get_dataset_most_frequent_motifs(trigram_plot_dat_presequences,
                                                         "cTP-mTP experimentally verified presequence",
                                                         n_motifs = 100)

most_frequent_SP <- get_dataset_most_frequent_motifs(trigram_plot_dat_presequences,
                                                     "SP experimentally verified presequence",
                                                     n_motifs = 100)

plot_most_frequent_motifs(trigram_plot_dat_presequences,
                          most_frequent_cTP,
                          colors2,
                          "100 most frequent motifs of cTP data set")

filter(trigram_plot_dat_presequences, motif %in% most_frequent_cTP) %>% 
  mutate(dataset = sapply(.[["dataset"]], function(i) strsplit(i, " ")[[1]][1])) %>% 
  pivot_wider(names_from = "dataset", values_from = "frequency") %>% 
  my_DT()


plot_most_frequent_motifs(trigram_plot_dat_presequences,
                          most_frequent_mTP,
                          colors2,
                          "100 most frequent motifs of mTP data set")

filter(trigram_plot_dat_presequences, motif %in% most_frequent_mTP) %>% 
  mutate(dataset = sapply(.[["dataset"]], function(i) strsplit(i, " ")[[1]][1])) %>% 
  pivot_wider(names_from = "dataset", values_from = "frequency") %>% 
  my_DT()


plot_most_frequent_motifs(trigram_plot_dat_presequences,
                          most_frequent_cTPmTP,
                          colors2,
                          "100 most frequent motifs of cTP-mTP data set")

filter(trigram_plot_dat_presequences, motif %in% most_frequent_cTPmTP) %>% 
  mutate(dataset = sapply(.[["dataset"]], function(i) strsplit(i, " ")[[1]][1])) %>% 
  pivot_wider(names_from = "dataset", values_from = "frequency") %>% 
  my_DT()


plot_most_frequent_motifs(trigram_plot_dat_presequences,
                          most_frequent_SP,
                          colors2,
                          "100 most frequent motifs of SP data set")

filter(trigram_plot_dat_presequences, motif %in% most_frequent_SP) %>% 
  mutate(dataset = sapply(.[["dataset"]], function(i) strsplit(i, " ")[[1]][1])) %>% 
  pivot_wider(names_from = "dataset", values_from = "frequency") %>% 
  my_DT()


frequent_for_all <- Reduce(intersect, list(most_frequent_cTP, most_frequent_cTPmTP, most_frequent_mTP, most_frequent_SP))
if(length(frequent_for_all) > 0) {
  plot_most_frequent_motifs(trigram_plot_dat_presequences,
                            frequent_for_all,
                            colors2,
                            "Intersection of the 100 most frequent motifs in all data sets")
}
```


### Position density of the 50 most frequent motifs for each dataset
```{r, fig.height = 20}
presequence_datasets_list <- datasets_list[c("cTP-mTP experimentally verified presequence",
                                             "cTP experimentally verified presequence",
                                             "mTP experimentally verified presequence",
                                             "SP experimentally verified presequence")]


most_frequent_cTP_25 <- get_dataset_most_frequent_motifs(trigram_plot_dat_presequences,
                                                         "cTP experimentally verified presequence",
                                                         n_motifs = 50)

most_frequent_mTP_25 <- get_dataset_most_frequent_motifs(trigram_plot_dat_presequences,
                                                         "mTP experimentally verified presequence",
                                                         n_motifs = 50)

most_frequent_cTPmTP_25 <- get_dataset_most_frequent_motifs(trigram_plot_dat_presequences,
                                                            "cTP-mTP experimentally verified presequence",
                                                            n_motifs = 50)

most_frequent_SP_25 <- get_dataset_most_frequent_motifs(trigram_plot_dat_presequences,
                                                        "SP experimentally verified presequence",
                                                        n_motifs = 50)

plot_motif_position_density(presequence_datasets_list, "cTP-mTP experimentally verified presequence",
                            locate_motifs(presequence_datasets_list["cTP-mTP experimentally verified presequence"], most_frequent_cTPmTP_25, params[["encoding"]]),
                            density_col = colors2[["cTP-mTP experimentally verified presequence"]])
plot_motif_position_density(presequence_datasets_list, "cTP experimentally verified presequence",
                            locate_motifs(presequence_datasets_list["cTP experimentally verified presequence"], most_frequent_cTP_25, params[["encoding"]]),
                            density_col = colors2[["cTP experimentally verified presequence"]])
plot_motif_position_density(presequence_datasets_list, "mTP experimentally verified presequence",
                            locate_motifs(presequence_datasets_list["mTP experimentally verified presequence"], most_frequent_mTP_25, params[["encoding"]]),
                            density_col = colors2[["mTP experimentally verified presequence"]])
plot_motif_position_density(presequence_datasets_list, "SP experimentally verified presequence",
                            locate_motifs(presequence_datasets_list["SP experimentally verified presequence"], most_frequent_SP_25, params[["encoding"]]),
                            density_col = colors2[["SP experimentally verified presequence"]])
```

### Motif positions within scaled sequences
```{r}
plot_motif_positions_scaled(presequence_datasets_list, "cTP-mTP experimentally verified presequence",
                            locate_motifs(presequence_datasets_list["cTP-mTP experimentally verified presequence"], most_frequent_cTPmTP_25, params[["encoding"]]),
                            type = "violin", density_col = colors2[["cTP-mTP experimentally verified presequence"]])
plot_motif_positions_scaled(presequence_datasets_list, "cTP experimentally verified presequence",
                            locate_motifs(presequence_datasets_list["cTP experimentally verified presequence"], most_frequent_cTP_25, params[["encoding"]]),
                            type = "violin", density_col = colors2[["cTP experimentally verified presequence"]])
plot_motif_positions_scaled(presequence_datasets_list, "mTP experimentally verified presequence",
                            locate_motifs(presequence_datasets_list["mTP experimentally verified presequence"], most_frequent_mTP_25, params[["encoding"]]),
                            type = "violin", density_col = colors2[["mTP experimentally verified presequence"]])
plot_motif_positions_scaled(presequence_datasets_list, "SP experimentally verified presequence",
                            locate_motifs(presequence_datasets_list["SP experimentally verified presequence"], most_frequent_SP_25, params[["encoding"]]),
                            type = "violin", density_col = colors2[["SP experimentally verified presequence"]])
```


## Motif presence/absence


### QuiPT most informative motifs

```{r}
binary_trigrams <- lapply(names(presequence_datasets_list), function(ith_dataset) {
  calculate_binary_ngram_matrix(presequence_datasets_list[[ith_dataset]], params[["k"]], params[["gaps"]]) %>%
    degenerate_ngrams(params[["encoding"]], binarize = TRUE) %>% 
    as.data.frame() %>% 
    mutate(dataset = ith_dataset)
}) %>% bind_rows()
binary_trigrams[is.na(binary_trigrams)] <- 0
```
cTP vs. mTP
```{r}
x <- select_quipt_informative_motifs(binary_trigrams,
                                     "cTP experimentally verified presequence",
                                     "mTP experimentally verified presequence")
print(paste0("Number of informative motifs: ", nrow(x)))
my_DT(x)
```
cTP-mTP vs. cTP & mTP
```{r}
x <- select_quipt_informative_motifs(binary_trigrams,
                                     "cTP-mTP experimentally verified presequence",
                                     c("cTP experimentally verified presequence", "mTP experimentally verified presequence"))
print(paste0("Number of informative motifs: ", nrow(x)))
my_DT(x)
```

cTP vs. cTP-mTP
```{r}
x <- select_quipt_informative_motifs(binary_trigrams,
                                     "cTP experimentally verified presequence",
                                     "cTP-mTP experimentally verified presequence")
print(paste0("Number of informative motifs: ", nrow(x)))
my_DT(x)
```

mTP vs. cTP-mTP
```{r}
x <- select_quipt_informative_motifs(binary_trigrams,
                                     "mTP experimentally verified presequence",
                                     "cTP-mTP experimentally verified presequence")
print(paste0("Number of informative motifs: ", nrow(x)))
my_DT(x)
```

SP vs. cTP & mTP
```{r}
x <- select_quipt_informative_motifs(binary_trigrams,
                                     "SP experimentally verified presequence",
                                     c("cTP experimentally verified presequence", "mTP experimentally verified presequence"))
print(paste0("Number of informative motifs: ", nrow(x)))
my_DT(x)
```


### FCBF most informative motifs
cTP vs. mTP
```{r}
y <- select_fcbf_informative_motifs(binary_trigrams,
                                    "cTP experimentally verified presequence",
                                    "mTP experimentally verified presequence")
print(paste0("Number of informative motifs: ", nrow(y)))
if(!is.null(y)) {
  if(nrow(y) > 1) {
      mutate(y, motif = gsub("X", "", motif)) %>% 
  my_DT()
  }
}
```

cTP-mTP vs. cTP & mTP
```{r}
y <- select_fcbf_informative_motifs(binary_trigrams,
                                    "cTP-mTP experimentally verified presequence",
                                    c("cTP experimentally verified presequence", "mTP experimentally verified presequence"))
print(paste0("Number of informative motifs: ", nrow(y)))
if(!is.null(y)) {
  if(nrow(y) > 1) {
      mutate(y, motif = gsub("X", "", motif)) %>% 
  my_DT()
  }
}
```

cTP vs. cTP-mTP
```{r}
y <- select_fcbf_informative_motifs(binary_trigrams,
                                    "cTP experimentally verified presequence",
                                    "cTP-mTP experimentally verified presequence")
print(paste0("Number of informative motifs: ", nrow(y)))
if(!is.null(y)) {
  if(nrow(y) > 1) {
      mutate(y, motif = gsub("X", "", motif)) %>% 
  my_DT()
  }
}
```

mTP vs. cTP-mTP
```{r}
y <- select_fcbf_informative_motifs(binary_trigrams,
                                    "mTP experimentally verified presequence",
                                    "cTP-mTP experimentally verified presequence")
print(paste0("Number of informative motifs: ", nrow(y)))
if(!is.null(y)) {
  if(nrow(y) > 1) {
      mutate(y, motif = gsub("X", "", motif)) %>% 
  my_DT()
  }
}
```

SP vs. cTP & mTP
```{r}
y <- select_fcbf_informative_motifs(binary_trigrams,
                                    "SP experimentally verified presequence",
                                    c("cTP experimentally verified presequence", "mTP experimentally verified presequence"))
print(paste0("Number of informative motifs: ", nrow(y)))
if(!is.null(y)) {
  if(nrow(y) > 1) {
      mutate(y, motif = gsub("X", "", motif)) %>% 
  my_DT()
  }
}
```
